MADA 8060E Project

### 2023-03-17 code/analysis_code/exploratory_analysis.qmd
Submit Part 3

---
title: "MADA Final Project"
subtitle: "Trends in New HIV Diagnoses in the US, 2008-2022"
author: Nicole Luisi
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

Note: There seem to be  lot of issues getting tables to render well in Word doc format, need to come back to this. I tried 5 or so different methods and they all look terrible when in a doc vs html. Also having issues with resolving crossrefs.

@@

# Summary/Abstract

The purpose of this analysis is to analyze US trends in new HIV diagnoses over the past 10 years. Previous research by [Chapin-Bardales et al. (2017)](https://doi.org/10.1016/j.annepidem.2017.04.002) evaluated trends in AIDS diagnosis rates by race/ethnicity between 1984-2013. This analysis will focus on new HIV diagnosesbetween  2008-2022; subgroups will include region, age group, gender, and race/ethnicity. [@Chapin2017]

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source

**Data Source**:

*Note: New diagnoses data are currently available through 2020. If 2021-2022 data are not available by the time the analysis needs to be completed, the analysis will be temporarily restricted to 2012-2020.*

+ HIV New Diagnoses: [AIDSVu](https://aidsvu.org/) [@Sullivan2020]

**Analysis**:

+ Rate (per 100,000) of new HIV diagnoses by age, gender, and race/ethnicity (national, regional)
+ Rate ratios for new HIV diagnoses by age, gender, and race/ethnicity (national, regional)
+ Estimated annual percent change (EAPC) for incidence
  - Calculated using [Joinpoint Regression Software (R-Callable Command-Line Version 4.9.0.0)](https://surveillance.cancer.gov/help/joinpoint/tech-help/citation)
+ Other modeling TBD based on course modules

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
Data files were downloaded from [AIDSVu](https://aidsvu.org/) and placed in the `data/raw_data/AIDSVu` subfolder. The new diagnosis files consist of one xlsx file per year for each level (national, regional, etc.).

## Data import and cleaning
The `processingcode.R` file includes the following steps:
+ Read in xlsx data files and combine all years for each level
+ Subset variables of interest and rename variables so all levels are consistent 
+ View/explore data
+ Calculate new rate ratio variables for subgroups 
+ Prepare combined national and regional file with all variables

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

@@

@tbl-natsummarytable shows a full summary of the National data.

```{r}
#| label: tbl-natsummarytable
#| tbl-cap: "National data summary table."
#| echo: FALSE
alldata<-readRDS(here("data", "processed_data", "allnewdx.rds"))
national<-alldata %>% filter(Geo %in% c("United States"))
national %>% gt::gt() %>% 
  gt::tab_header(title = "National Data") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))
```
@fig-natplots shows exploratory plots for the national rates by subgroup.

```{r}
#| label: fig-natplots
#| fig-cap: "National rates by subgroup"
#| echo: FALSE
national<-alldata %>% filter(Geo %in% c("United States"))
natfig <- plot_ly(national, x = ~Year, y = ~Overall_Rate, name = "Overall", type="scatter", mode = "lines", line = list(width = 3, dash = "dash", color = "gray")) 
natfig <- natfig %>% add_trace(y = ~Male_Rate, name = "Male", mode = 'lines', line = list(width = 3, dash = "solid", color = "blue"))
natfig <- natfig %>% add_trace(y = ~Female_Rate, name = "Female", mode = 'lines', line = list(width = 3, dash = "solid", color = "green"))
natfig %>% layout(title="National New Diagnosis Rate by Sex", legend=list(title=list(text='Sex')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate per 100,000'))

@@

@tbl-natratios shows a summary of selected rate ratios for race/ethnicity and sex.

```{r}
#| label: tbl-natratios
#| tbl-cap: "National rate ratios by select subgroups"
#| echo: FALSE
national<-alldata %>% filter(Geo %in% c("United States"))
national_sub <- national %>% dplyr::select("Year", "Black_Rate", "White_Rate", "Hispanic_Rate", "black_white_rateratio", "hispanic_white_rateratio", "Male_Rate", "Female_Rate", "female_male_rateratio")
national_sub %>% gt::gt() %>% 
  gt::tab_header(title = "Rate Ratios") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))%>%
  gt::tab_style(style = list(gt::cell_fill(color = "#F1EB9C"),
                             gt::cell_text(weight = "bold")), 
                locations = gt::cells_body(columns = c(black_white_rateratio,hispanic_white_rateratio,female_male_rateratio)))
```

@tbl-regsummarytable shows a full summary of the Regional data.

```{r}
#| label: tbl-regsummarytable
#| tbl-cap: "Regional data summary table."
#| echo: FALSE
alldata<-readRDS(here("data", "processed_data", "allnewdx.rds"))
regional<-alldata %>% filter(Geo %in% c("Northeast", "Midwest", "South", "West"))
regional %>% gt::gt() %>% 
  gt::tab_header(title = "Regional Data") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))
```

@fig-regplots shows an exploratory plot for regional overall new diagnosis rates with female-male rate ratios.

```{r}
#| label: fig-regplots
#| tbl-cap: "Regional plots"
#| echo: FALSE
alldata<-readRDS(here("data", "processed_data", "allnewdx.rds"))
regional<-alldata %>% filter(Geo %in% c("Northeast", "Midwest", "South", "West"))
ggplot(regional, aes(x=Year, y=log(Overall_Rate), size=log(female_male_rateratio), color=Geo)) +
    geom_point(alpha=0.7) + scale_x_continuous(breaks=c(2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  ggtitle("Regional New Diagnosis Rates with Female-Male Rate Ratios") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Year") + ylab("Log of Overall NewDx Rate")
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by sex."
#| echo: FALSE
#knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","resulttable2.rds"))
#knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [] discusses types of analyses.

These papers [] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

>>>>

---
title: "Exploratory analysis script"
author: "Nicole Luisi"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

@@

## Load data

```{r}
alldata<-readRDS(here("data", "processed_data", "allnewdx.rds"))
```

@@

Full summary of the National data.

```{r}
national<-alldata %>% filter(Geo %in% c("United States"))
national %>% gt::gt() %>% 
  gt::tab_header(title = "National Data") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))
```

Exploratory plots for the national rates by subgroup.

```{r}
national<-alldata %>% filter(Geo %in% c("United States"))
natfig <- plot_ly(national, x = ~Year, y = ~Overall_Rate, name = "Overall", type="scatter", mode = "lines", line = list(width = 3, dash = "dash", color = "gray")) 
natfig <- natfig %>% add_trace(y = ~Male_Rate, name = "Male", mode = 'lines', line = list(width = 3, dash = "solid", color = "blue"))
natfig <- natfig %>% add_trace(y = ~Female_Rate, name = "Female", mode = 'lines', line = list(width = 3, dash = "solid", color = "green"))
natfig %>% layout(title="National New Diagnosis Rate by Sex", legend=list(title=list(text='Sex')), 
                  xaxis = list(title = 'Year'),
                  yaxis = list(title = 'Rate per 100,000'))

@@

Summary of selected rate ratios for race/ethnicity and sex.

```{r}
national<-alldata %>% filter(Geo %in% c("United States"))
national_sub <- national %>% dplyr::select("Year", "Black_Rate", "White_Rate", "Hispanic_Rate", "black_white_rateratio", "hispanic_white_rateratio", "Male_Rate", "Female_Rate", "female_male_rateratio")
national_sub %>% gt::gt() %>% 
  gt::tab_header(title = "Rate Ratios") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))%>%
  gt::tab_style(style = list(gt::cell_fill(color = "#F1EB9C"),
                             gt::cell_text(weight = "bold")), 
                locations = gt::cells_body(columns = c(black_white_rateratio,hispanic_white_rateratio,female_male_rateratio)))
```

Full summary of the Regional data.

```{r}
regional<-alldata %>% filter(Geo %in% c("Northeast", "Midwest", "South", "West"))
regional %>% gt::gt() %>% 
  gt::tab_header(title = "Regional Data") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = Year))
```

Exploratory plot for regional overall new diagnosis rates with female-male rate ratios.

```{r}
regional<-alldata %>% filter(Geo %in% c("Northeast", "Midwest", "South", "West"))
ggplot(regional, aes(x=Year, y=log(Overall_Rate), size=log(female_male_rateratio), color=Geo)) +
    geom_point(alpha=0.7) + scale_x_continuous(breaks=c(2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  ggtitle("Regional New Diagnosis Rates with Female-Male Rate Ratios") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Year") + ylab("Log of Overall NewDx Rate")
```

### 2023-03-17 products/manuscript/Manuscript.qmd
Submit Part 3

---
title: "MADA Final Project"
subtitle: "Trends in PrEP Use in the US, 2012-2022"
author: Nicole Luisi
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

# PART 1

**Overview**:

The purpose of this analysis is to measure the public health impact of pre-exposure prophylaxis (PrEP) in the US over the past 10 years. Previous research by [Sullivan et al. (2018)](https://doi.org/10.1016/j.annepidem.2018.06.009) evaluated trends in the prevalence of PrEP use from 2012-2017. Outcomes include period prevalence of PrEP use and PrEP-to-need ratio (PnR) by year. This analysis will extend the previous work to include data through the year 2022; subgroups will include region, state, age group, gender, and race/ethnicity *(race/ethnicity PrEP data were not available at the time of the previous analysis and were not included)*. [@SULLIVAN2018833]

*Note: New diagnoses data are currently available through 2020; PrEP data through 2021. If 2021-2022 data are not available by the time the analysis needs to be completed, the analysis will be temporarily restricted to 2012-2020.*

+ HIV New Diagnoses: [AIDSVu](https://aidsvu.org/) [@Sullivan2020] 
+ PrEP and PnR: IQVIA (Quintiles IMS Holdings)
+ Annual Population Estimates: National Historical Geographic Information System (NHGIS), American Community Surveys (ACS), US Census Bureau

+ N/% of annual PrEP users by age, gender, race/ethnicity, and region
+ Annual prevalence of PrEP use per 100,000 population by age, gender, race/ethnicity, region, and state
+ Estimated annual percent change (EAPC) for prevalence
  - Calculated using [Joinpoint Regression Software (R-Callable Command-Line Version 4.9.0.0)](https://surveillance.cancer.gov/help/joinpoint/tech-help/citation)
+ Prep-to-need ratio by age, gender, race/ethnicity, and region
+ Other modeling TBD based on course modules

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
#library(here)
#library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

@@

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in []

@@

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

@@

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
#resulttable=readRDS("../../results/summarytable.rds")
#knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by sex."
#| echo: FALSE
#knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","resulttable2.rds"))
#knitr::kable(resulttable2)
```

@@

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [] discusses types of analyses.

These papers [] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

@@

# References

>>>>

---
title: "MADA 8060E Project"
subtitle: "Trends in New HIV Diagnoses in the US, 2008-2022"
author: Nicole Luisi
date: "`r Sys.Date()`"
format:
  html:
    self-contained: true
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# load R packages
library(here) 
library(ggplot2)
library(skimr) 
library(plotly)
library(gt)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
library(gridExtra)
library(ggpubr)
```

# Summary/Abstract

The purpose of this analysis is to analyze US trends in new HIV diagnoses over the past 10 years. Previous research by [Chapin-Bardales et al. (2017)](https://doi.org/10.1016/j.annepidem.2017.04.002) evaluated trends in AIDS diagnosis rates by race/ethnicity between 1984-2013. This analysis will focus on new HIV diagnoses between 2008-2022; subgroups will include region, age group, gender, and race/ethnicity. [@Chapin2017]

@@

## Description of data and data source

*Note: New diagnoses data are currently only available through 2020 on AIDSVu and CDC Atlas. If 2021-2022 data are not available by the time the analysis needs to be completed, the analysis will be temporarily restricted to 2012-2020.*

* HIV New Diagnoses: [AIDSVu](https://aidsvu.org/) [@Sullivan2020]

* Rate (per 100,000) of new HIV diagnoses by age, gender, and race/ethnicity (national, regional)
* Rate ratios for new HIV diagnoses by age, gender, and race/ethnicity (national, regional)
* Average annual percent change (AAPC) for incidence (calculated using [Joinpoint Regression Software (Version 4.9.1.0)](https://surveillance.cancer.gov/help/joinpoint/tech-help/citation))

## Questions/Hypotheses to be addressed
This analysis will evaluate national and regional trends in HIV new diagnoses between 2008-2020 by sex, age, and race/ethnicity.

@@

## Data aquisition
Data files were downloaded from [AIDSVu](https://aidsvu.org/) and placed in `data/raw_data/AIDSVu` sub-folders. The new diagnosis files consist of one xlsx file per year for each level (national, regional, etc.).

## Data import and cleaning
The `processingcode.R` file includes the following steps:

* Read in xlsx data files and combine all years for each level
* Subset variables of interest and rename variables so all levels are consistent 
* View/explore data
* Calculate new rate ratio variables for subgroups (White=ref)
* Prepare combined national and regional file with all variables

The `analysis.qmd` file includes the following steps:

* Read in result data from external Joinpoint analyses
* Clean and combine Joinpoint analysis results
* View/explore data
* Create tables and plots of results

## Statistical analysis
Rate ratios for race/ethnicity were calculated by dividing the new diagnosis rate for each group by the new diagnosis rate for the reference group (non-Hispanic White).

[Joinpoint Regression Software (Version 4.9.1.0)](https://surveillance.cancer.gov/help/joinpoint/tech-help/citation) was used to test for significant joinpoints (years) to identify  trends in new diagnoses rates, nationally and regionally, by sex, age, and race/ethnicity. Trends for rate ratios were also used to assess disparities by race/ethnicity over time. For Joinpoint analyses, the maximum number of joinpoints was set to 3, and standard errors were auto-calculated by the tool under the heteroscedasticity assumption. The average annual percent changes (AAPCs) from 2008-2020 were also calculated for all subsets.

@@

## Descriptive analysis

@fig-natplotsex shows line plot for national new diagnosis rates by sex.

```{r}
#| label: fig-natplotsex
#| fig-cap: "National new diagnosis rates by sex"
#| echo: FALSE
knitr::include_graphics(here("results","US_rates_bysex.png"))
```

@fig-natplotage shows line plot for national new diagnosis rates by age group.

```{r}
#| label: fig-natplotage
#| fig-cap: "National new diagnosis rates by age group"
#| echo: FALSE
knitr::include_graphics(here("results","US_rates_byage.png"))
```

@fig-natplotrace shows line plot for national new diagnosis rates by race/ethnicity.

```{r}
#| label: fig-natplotrace
#| fig-cap: "National new diagnosis rates by race/ethnicity"
#| echo: FALSE
knitr::include_graphics(here("results","US_rates_byrace.png"))
```

Trends by sex and age group are as expected. For sex and age, all rates have declined between 2008-2020 and groups with the highest incidence had the steepest declines over time, particularly after 2016. The remainder of this analysis will focus on trends and disparities by race/ethnicity.

@fig-regplotrace shows a bubble plot comparing regional new diagnosis (log) rates by race/ethnicity. The color represents race/ethnicity and the size of the point represents the log of the new diagnosis rate.

```{r}
#| label: fig-regplotrace
#| fig-cap: "Regional new diagnosis rate (log) by race/ethnicity"
#| echo: FALSE
knitr::include_graphics(here("results","racerates_region.png"))
```

@fig-natplotraceratios shows national plot of log race ratios for new diagnoses (White=ref).

```{r}
#| label: fig-natplotraceratios
#| fig-cap: "National race ratios"
#| echo: FALSE
knitr::include_graphics(here("results","racerates_allUS.png"))
```

@fig-natplotdiffbwh shows national plot of White-Black and White-Hispanic rate differences by year.

```{r}
#| label: fig-natplotdiffbwh
#| fig-cap: "White-Black and White-Hispanic differences in national new diagnosis rate"
#| echo: FALSE
knitr::include_graphics(here("results","nat_diff_bwh.png"))
```

@tbl-allrates shows a full summary of new diagnosis rates for all US by group.

```{r}
#| label: tbl-allrates
#| tbl-cap: "All US New diagnosis Rates"
#| echo: FALSE
resulttable=readRDS(here("results","all_us_rates.rds"))
kable(resulttable,caption = "All US New Diagnosis Rates", col.names=c("Year", "Overall", "Male", "Female", "Black", "White", "Hispanic", "Asian", "American Indian/Alaska Native", "Multiracial", "Native Hawaiian/Pacific Islander", "Age 13-24", "Age 25-34", "Age 35-44", "Age 45-54", "Age 55+"), align="rc") %>% 
  row_spec(row = 0, color = "white", background = "#004987") %>% kable_styling(full_width=T, bootstrap_options = "striped", font_size=10)  
```

## Statistical analysis

@tbl-allaapc shows a summary of average annual percent change (AAPC) for each subgroup and geographic region.

```{r}
#| label: tbl-allaapc
#| tbl-cap: "All AAPCs"
#| echo: FALSE
resulttable=readRDS(here("results","aapc_all_results.rds"))
kable(resulttable,caption = "All AAPC Results", 
      col.names=c("Stratification", "Geographic Area", "AAPC", "AAPC_lCI", "AAPC_uCI", "p-value"),      align="rc") %>% 
  row_spec(row = 0, color = "white", background = "#004987") %>% kable_styling(full_width=T, bootstrap_options = "striped", font_size=10) %>% 
  pack_rows("Overall", 1, 5) %>%
  pack_rows("Female", 6, 10) %>%
  pack_rows("Male", 11, 15) %>%
  pack_rows("Age 13-24", 16, 20) %>%
  pack_rows("Age 25-34", 21, 25) %>%
  pack_rows("Age 35-44", 26, 30)  %>%
  pack_rows("Age 45-54", 31, 35)  %>%
  pack_rows("Age 55+", 36, 40)  %>%
  pack_rows("American Indian/Alaska Native", 41, 44) %>%
  pack_rows("Asian", 45, 49) %>%
  pack_rows("Black", 50, 54) %>%
  pack_rows("Hispanic", 55, 59) %>%
  pack_rows("Multiracial", 60, 64) %>%
  pack_rows("Native Hawaiian/Pacific Islander", 65, 67) %>%
  pack_rows("White", 68, 72) %>%
  pack_rows("AIAN-White", 73, 76) %>%
  pack_rows("Asian-White", 77, 81) %>%
  pack_rows("Black-White", 82, 86) %>%
  pack_rows("Hispanic-White", 87, 91) %>%
  pack_rows("Multi-White", 92, 96) %>%
  pack_rows("NHPI-White", 97, 99) 
```

@tbl-allapc shows a full summary of annual percent change (APC) for each segment of time within each subgroup and geographic region.

```{r}
#| label: tbl-allapc
#| tbl-cap: "All APCs"
#| echo: FALSE
resulttable=readRDS(here("results","apc_all_results.rds"))
kable(resulttable,caption = "All APC Results", align="rc") %>% 
  row_spec(row = 0, color = "white", background = "#004987") %>% kable_styling(full_width=T, bootstrap_options = "striped", font_size=10) %>% 
  pack_rows("Overall", 1, 15) %>%
  pack_rows("Female", 16, 24) %>%
  pack_rows("Male", 25, 32) %>%
  pack_rows("Age 13-24", 33, 41) %>%
  pack_rows("Age 25-34", 42, 54) %>%
  pack_rows("Age 35-44", 55, 63)  %>%
  pack_rows("Age 45-54", 64, 70)  %>%
  pack_rows("Age 55+", 71, 76)  %>%
  pack_rows("American Indian/Alaska Native", 77, 80) %>%
  pack_rows("Asian", 81, 90) %>%
  pack_rows("Black", 91, 102) %>%
  pack_rows("Hispanic", 103, 113) %>%
  pack_rows("Multiracial", 114, 120) %>%
  pack_rows("Native Hawaiian/Pacific Islander", 121, 123) %>%
  pack_rows("White", 124, 130) %>%
  pack_rows("AIAN-White", 131, 134) %>%
  pack_rows("Asian-White", 135, 142) %>%
  pack_rows("Black-White", 143, 147) %>%
  pack_rows("Hispanic-White", 148, 156) %>%
  pack_rows("Multi-White", 157, 165) %>%
  pack_rows("NHPI-White", 166, 168) 
```

@fig-natplotbwhjpoint shows national plot of new diagnosis Joinpoints for race/ethnicity (Black, White, Hispanic).

```{r}
#| label: fig-natplotbwhjpoint
#| fig-cap: "National Joinpoint results comparing Black, White, Hispanic"
#| echo: FALSE
knitr::include_graphics(here("results","raceBWH_joinpoints_allUS.png"))
```

@fig-natplotraceratioaapc shows regional plot of new diagnosis AAPCs for race/ethnicity.

```{r}
#| label: fig-natplotraceratioaapc
#| fig-cap: "Regional Joinpoint AAPC results by race/ethnicity"
#| echo: FALSE
knitr::include_graphics(here("results","raceratio_byregion.png"))
```

@@

@@

## References
